---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->
<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental) [![CRAN status](https://www.r-pkg.org/badges/version/nwslR)](https://CRAN.R-project.org/package=nwslR) [![Travis build status](https://travis-ci.org/adror1/nwslR.svg?branch=master)](https://travis-ci.org/adror1/nwslR) <!-- badges: end -->


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# nwslR


An R package providing users with NWSL player, team, franchise, and stadium data, as well as functionality to glean further data and information. 

## Install 

You can install the current version of nwslR from GitHub using:

```{r}
devtools::install_github("adror1/nwslR")
```

## Load

This is a basic example which shows you how to solve a common problem:

```{r}
library(nwslR)
```

## Available Tables

`fp_overall_season_stats`: season-by-season data for each NWSL field player

`goalkeeper_season_stats`: season-by-season data for each NWSL goalkeeper

`team_stats_season`: season level statistics for each team from the league's inception through present-day

`franchise`: team names, ids, and locations throughout their history in the league

`stadium`: historical and current stadium information and statistics

## Example: Player Statistics

```{r}
nwslR::fieldplayer_overall_season_stats
```

## Example: Team IDs
```{r}
nwslR::franchise
```

## Functionality
```{r, eval=FALSE}
nwslR::get_adv_team_stats() #scrapes and formats hundreds of advanced team statistics from the NWSL website
nwslR::get_adv_player_stats() #scrapes and formats advanced individual player stats from the NWSL website
nwslR::player_search() #searches data tables based on player information such as name, nation, or position
```

You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date.

You can also embed plots, for example:

In that case, don't forget to commit and push the resulting figure files, so they display on GitHub!
